On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-01 tt_theorem.tex
12/1 before lunch

\begin{lemma}
\label{TTT_entrance_Tab}
Let $\Xi$ be a deterministic unary oritatami system at delay $\delta = 1$, arity $\alpha =2$. 
Assume $\Xi$ stabilizes the transcript until $w[i-1]$. If $w[i]$ is stabilized at an entrance point of tunnel $A$ or $B$, then $\#bc(C_{i-1}) > \#bc(C_{i})$.
\end{lemma}

Let us first consider cases of $\delta \geq 3, \alpha = 1$. These cases are clearly true because of Lemma~\ref{TTT_exit}.

%\begin{figure}
%  \begin{center}
%    \begin{tikzpicture}
%      \draw[thick, gray]
%      (0.5,1)--(0,1)--(0,4)--(4,4)--(4,1)--(3.5,1);
%      \node at(2,1) {Outside};
%      \node at (2,3.2) [rectangle, draw, rounded corners] (enter) {Entrance of a tunnel};
%      \node at(2,1.8) [rectangle, draw, rounded corners] (outside) {Otherwise};
%
%      
%      
%      \draw[thick, gray]
%      (7.5,1)--(7,1)--(7,4)--(11,4)--(11,1)--(10.5,1);
%      \node at(9,1) {Inside of a tunnel};
%      \node at(9,3.2) [rectangle, draw, rounded corners] (next-tunnel) {Successor in inside};
%      \node at(9,1.8) [rectangle, draw, rounded corners] (next-outside) {Successor in outside};
%
%      \draw[->]
%      (outside) edge [bend left, blue] node[left] {$s=b,\Delta=-1$} (enter)
%                   edge [loop left, blue] node[below left] {$s=b,\Delta=\pm 0$} (outside)
%      (enter)   edge [bend left, blue] node[above] {$s=b,\Delta=-1$} (next-tunnel)
%                edge [bend left, blue] node[above] {$s=b,\Delta=-a$} (next-outside)
%                edge [bend left, blue] node[right] {$s=b,\Delta=\pm 0$}  (outside)
%      (next-tunnel) edge[teal]         node[right] {$s=t,\Delta=-a$} (next-outside)
%                edge [loop right, teal]node[right] {$s=t,\Delta=\pm 0$}    (next-tunnel)
%      (next-outside) edge[bend left, teal] node[above] {$s=t,\Delta=+a$}  (outside)
%                edge [bend left, teal]     node[above] {$s=t,\Delta=+(a-1)$}(enter)
%      ;
%      
%    \end{tikzpicture}
%    \caption{Increment on Tunnel A,B}
%    \label{TTT_position}
%  \end{center}
%\end{figure}

%\begin{figure}
%  \begin{center}
%    \begin{tikzpicture}
%      \node at (0,4)   [rectangle, draw,rounded corners] (outside1) {Outside of tunnels};
%      \node at (3.5,4) [rectangle, draw,rounded corners] (enter1)   {Entrance of a tunnel};
%      \node at (7,4)   [rectangle, draw, rounded corners] (inside1) {Inside of a tunnel C};
%      \node at (11,4)  [rectangle, draw, rounded corners] (exit1)   {Inside of a tunnel A or B};
%      \draw[->] (outside1)--(enter1);
%      \draw[->] (enter1)  --(inside1);
%      \draw[->] (inside1) --(exit1);
%      
%      
%      \node at (0,3)   [rectangle, draw,rounded corners] (outside2) {Inside of a tunnel};
%      \node at (3.5,3) [rectangle, draw,rounded corners] (enter2)   {Entrance of a tunnel};
%      \node at (7,3)   [rectangle, draw, rounded corners] (inside2) {Inside of a tunnel C};
%      \node at (11,3)  [rectangle, draw, rounded corners] (exit2)   {Inside of a tunnel A or B};     
%      \draw[->] (outside2)--(enter2);
%      \draw[->] (enter2)  --(inside2);
%      \draw[->] (inside2) --(exit2);
%
%      \node at (0,1)   [rectangle, draw,rounded corners] (outside3) {Outside of tunnels};
%      \node at (3.5,1) [rectangle, draw,rounded corners] (enter3)   {Entrance of a tunnel};
%      \node at (7,1)   [rectangle, draw, rounded corners] (inside3) {Inside of a tunnel C};
%      \node at (11,1)  [rectangle, draw, rounded corners] (exit3)   {Outside of tunnels};     
%      \draw[->] (outside3)--(enter3);
%      \draw[->] (enter3)  --(inside3);
%      \draw[->] (inside3) --(exit3);
%
%      \node at (0,0)   [rectangle, draw,rounded corners] (outside4) {Inside of atunnel};
%      \node at (3.5,0) [rectangle, draw,rounded corners] (enter4)   {Entrance of a tunnel};
%      \node at (7,0)   [rectangle, draw, rounded corners] (inside4) {Inside of a tunnel C};
%      \node at (11,0)  [rectangle, draw, rounded corners] (exit4)   {Outside of tunnels};     
%      \draw[->] (outside4)--(enter4);
%      \draw[->] (enter4)  --(inside4);
%      \draw[->] (inside4) --(exit4);
%    \end{tikzpicture}
%    \caption{Case of Tunnel C}
%    \label{TTT_tunnelC_graph}
%  \end{center}
%\end{figure}

@@

            \fill[blue] (0,0) circle [radius=0.05];
            \node[below] at (0,0) {$p^\prime$};

@@

            \fill(0,0) circle [radius=0.05];
            \node[below] at (0,0) {$p^\prime$};

@@

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$p^\prime$};

@@

\begin{paragraph}{}
$\delta = 1, \alpha = 2$\\
Let $\Xi$ be a unary oritatami system at $\delta = 1, \alpha = 2$. We assume $S[h..i+1] = bt^{(i-h)}b \quad (h<i)$ and
let $a$ be $\#bc(C_{i-2}) - \#bc(C_{i-1}) = a$. Then, $\#bc(C_{i}) - \#bc(C_{i-1}) \leq a$ as follows. Also, if $i - h > 1$ and $j$ is such that $h<j<i$, then $\#bc(C_{j-1}) \geq \#bc(C_{j})$ because all neighbors of $w[j]$ are occupied by beads forming the tunnel so that any $w[i+1..]$ cannot reach neighbors of $w[j]$. Thus, $\#bc(C_{h-1}) \geq \#bc(C_i)$ and $\#bc(C_{i-2}) \geq \#bc(C_i)$.
\end{paragraph}

@@

\begin{paragraph}{}
$\delta = 1, \alpha \geq 3$\\
Let $\Xi$ be a unary oritatami system at $\delta = 1, \alpha \geq 3$. We assume $w[i]$ is stabilized at exit of tunnel.
In all cases $\#bc(C_{i-1}) > \#bc(C_{i})$. Moreover, if $S[k] = t (k \leq i)$, then $\#bc(C_{k-1}) > \#bc(C_{k})$ because both sides of the path $p$ in Fig.\ref{TTT_inside_tunnel} ($n_1, n_2$) have two free points and one of $n_1, n_2$ is not the predecessor so that it has hand and moreover that $w[k]$ supplies any binding capabilities because its neighbors are occupied by beads of tunnel.
\end{paragraph}

\begin{itemize}
\item{Case of $n_1$ and $n_2$ are free}\\
  In $\alpha \geq 3$, if three neighbors of a bead leave, then it can supply two binding capabilities. Therefore predecessor $n_5$ has to be bound $n_3$ and $n_4$, and $p^\prime$, too. In this case, at least four bindings are consumed and at most two are added. Thus, it consumes some binding capabilities, overall.

\item{Case of $n_1$ is occupied}\\
  In this case, $n_4$ leave at least two bindings and $n_3$, $n_1$ also leave at least one binding. Therefore $n_5$ has to be bound $n_3$ and $n_4$, and $p^\prime$ also has to be bound $n_1$ and $n_4$. In this case, at least four bindings are consumed and at most two are added. Thus, it consumes some binding capabilities, totally.
\end{itemize}

\begin{figure}
  \centering
    \begin{tabular}{cc}
    \begin{minipage}{0.66\hsize}
    \centering
      \begin{tabular}{cc}
      \begin{minipage}{0.45\linewidth}
      \centering
        \begin{tikzpicture}
	\draw[white] (0:0) -- (120:1);
	\draw[white] (0:0) -- (-60:1);
	\draw[white] (0:0) -- (0:1);

            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$p^\prime$};

@@

@@

            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$p^\prime$};

@@

            \node[transform canvas={shift=(0:1)},right] {$n_0$};
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

          \node at (0,-2) {Point $n_1$ is occupied};
        \end{tikzpicture}
      \end{minipage}
    \end{tabular}
    \caption{Exit of Tunnel}
    \label{TTT_tunnel_exit}
    \end{minipage}

    \begin{minipage}{0.33\hsize}
    \centering
     \begin{tikzpicture}
	\draw[white] (0:0) -- (120:1);
	\draw[white] (0:0) -- (-60:1);
            \fill[blue](0,0) circle [radius=0.05];
            \draw[blue] (0:1) circle [radius = 0.05];
            \draw[->] (0:0.1) -- node[above]{$p$} (0:0.9);

            \foreach \theta in {60, -60, 120,-120,180}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
                \end{tikzpicture}

    \caption{Inside tunnel}
    \label{TTT_inside_tunnel}
    \end{minipage}

    \end{tabular}
\end{figure}
\end{proof}

@@

>>>>

\begin{lemma}
\label{TTT_entrance_Tab}
Let $\Xi$ be a deterministic unary oritatami system at delay $\delta = 1$, arity $\alpha =2$. 
Assume $\Xi$ stabilizes the transcript until $w[i-1]$. If $S[i] = b$ and $S[i+1] \in \{ t_s, t_o \} $, then $\#bc(C_{i-1}) > \#bc(C_{i})$.
\end{lemma}

@@

            \fill[blue] (0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};

@@

            \fill(0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};

@@

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
            \fill[blue](0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};

@@

Let $\Xi$ be a unary oritatami system at $\delta = 1, \alpha = 2$. We assume $S[h..i+1] = bt^{(i-h)}b \quad (h<i)$ and
let $a$ be $\#bc(C_{i-2}) - \#bc(C_{i-1}) = a$. Then, $\#bc(C_{i}) - \#bc(C_{i-1}) \leq a$ as follows. Also, if $i - h > 1$ and $j$ is such that $h<j<i$, then $\#bc(C_{j-1}) \geq \#bc(C_{j})$ because all neighbors of $w[j]$ are occupied by beads forming the tunnel so that any $w[i+1..]$ cannot reach neighbors of $w[j]$. Thus, $\#bc(C_{h-1}) \geq \#bc(C_i)$ and $\#bc(C_{i-2}) \geq \#bc(C_i)$.

@@

\begin{figure}
  \centering
    \begin{tabular}{cc}
    \centering
      \begin{minipage}{0.45\linewidth}
      \centering
        \begin{tikzpicture}
	\draw[white] (0:0) -- (120:1);
	\draw[white] (0:0) -- (-60:1);
	\draw[white] (0:0) -- (0:1);

            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$w[i]$};

@@

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$w[i-1]$};

@@

            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$w[i]$};

@@

            \node[transform canvas={shift=(0:1)},right] {$n_0$};
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},above] {$w[i-1]$};

          \node at (0,-2) {Point $n_1$ is occupied};
        \end{tikzpicture}
      \end{minipage}
    \end{tabular}
    \caption{Exit of Tunnel}
    \label{TTT_tunnel_exit}
\end{figure}
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%%%
%%%%%%
%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let us first consider cases of $\delta \geq 3, \alpha = 1$. 
See Fig.~\ref{TTT_tunnel_exit}. Consider the stabilization of $w[i]$. 
This bead $w[i]$, once stabilized, shares two neighbors with its predecessor $w[i-1]$, which are denoted by $n_3, n_4$. 
Both of them have been already occupied because $S[i] = t$.

Since $S[i+1] \neq \blacksquare$, at least one of the other three neighbors, denoted by $n_0, n_1, n_2$, must be free. 
Assume that in the neighborhood of $w[i]$, there are two beads with one free neighbor even after $w[i]$ is stabilized. 
Before the stabilization of $w[i]$, such a bead had two free neighbors, and hence, is provided with at least one free bond. 
Thus, $w[i]$ is to be bonded to these two beads, and it decreases the binding capability by at least 1. 
It now suffices to check that this assumption holds no matter how $n_0, n_1, n_2$ are occupied as long as at least one of them is left free.
